---
title: "oame"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Ocean Account - Maine data
The [oame R package]() serves Maine-centric ocean accounts data.


## Requirements

  + [R v4.2+](https://www.r-project.org/)
  + [rlang](https://CRAN.R-project.org/package=rlang)
  + [ggplot2](https://CRAN.R-project.org/package=ggplot2)
  + [dplyr](https://CRAN.R-project.org/package=dplyr)
  + [tidyr](https://CRAN.R-project.org/package=tidyr)
  + [readr](https://CRAN.R-project.org/package=readr)
  + [yaml](https://CRAN.R-project.org/package=yaml)


## Clone the repos

Clone the original (or a fork of the orginal) to your local machine/

## Set up the data path

TO keep things simple for the end user, set up a text file (YAML format) in the users home directory: `~/.oame`

The from within R store the correct data path into the file.  Here's an example, but obviously you would amend the path to fit your situation.

```
library(yaml)
x = list(data_path = "/mnt/ecocast/corecode/R/oceanaccounts/oame/data" )
write_yaml(x, "~/.oame")
```

## Installation

```
remotes::install_github("BigelowLab/ocean-accounts-maine/package", upgrade = FALSE)
```

## Reading data


```{r nao}
library(oame)
library(dplyr)

nao = read_nao() |> glimpse()
```

```{r amo}
amo = read_amo() |> glimpse()
```

```{r landings}
landings = read_dmr_landings("merged") |> glimpse()
```


## Plotting

The NAO and AMO data are identical in structure, so they share three plotting schemes... but each has a wrapper of it's own: `plot_nao()` and `plot_amo()`

```{r amo_timeseries}
plot_amo(amo, type = "timeseries")
```

```{r amo_monthly}
plot_amo(amo, type = "monthly")
```

```{r amo_climatology}
plot_amo(amo, type = "climatology")
```


```{r chloropleth}
map_species_by_county()
```

```{r cartogram}
map_species_by_county(spp = "Lobster American", style = "cartogram")
```


